---
title: "Question2"
author: "Yuming Shao"
date: "2018/2/15"
output: pdf_document
---

Question 2.

a.
Generate X and Y for case1.
```{r, include=FALSE}
sigma= matrix(1,50,50)
for (i in 1:50){
  for (j in 1:50){
    sigma[i,j]=2*0.5^abs(i-j)
    }
}
#case 1
library("MASS")
x1<-mvrnorm(n=2^16, rep(0,50), sigma)
beta<-c(rep(1,10),rep(0.11,30),rep(1,10))
y1<-x1%*%beta+rnorm(n=2^16,mean=0,sd=9)
```

Generate X and Y for case2.
```{r}
sigma= matrix(1,50,50)
for (i in 1:50){
  for (j in 1:50){
    sigma[i,j]=2*0.5^abs(i-j)
    }
}
#case 2
library("mvtnorm")
library(R.matlab)
x2<-rmvt(n=2^16,sigma,df=1 ,delta=rep(0,50))
beta<-c(rep(1,10),rep(0.11,30),rep(1,10))
y2<-x2%*%beta+rnorm(n=2^16,mean=0,sd=9)
```

Caculate $MSE_{\beta}$,$Bias_{\beta}$,$Variance_{\beta}$ for case1:
```{r}
zz <- cbind(x1, y1)
zz<-data.frame(zz)

N <- nrow(x1)
C <- 1000 
S <- 1000
sumb2 <- rep(0,50)
for (i in 1:C){   # a loop over the number of subsamples
  set.seed(3*i)   # a different seed for each subsample  
  subsample <- zz[sample(1:N, size=S, replace=TRUE), ]
  mod2 <- lm(X51~X1:X50,data=subsample)
  #sum b2 for all subsamples:
  sumb2 <- sumb2 + coef(mod2)[[2]]
}
betahat<-sumb2/C
#MSE
mean((betahat-beta)^2)

#Bias
sum(betahat-beta)

#Variance
var(betahat)
```


Caculate $MSE_{\beta}$,$Bias_{\beta}$,$Variance_{\beta}$ for case2:
```{r}
#Case 2
zz2 <- cbind(x2, y2)
zz2<-data.frame(zz2)

N <- nrow(x2)
C <- 1000 
S <- 1000
sumb2 <- rep(0,50)
for (i in 1:C){   # a loop over the number of subsamples
  set.seed(3*i)   # a different seed for each subsample  
  subsample <- zz2[sample(1:N, size=S, replace=TRUE), ]
  mod2 <- lm(X51~X1:X50,data=subsample)
  #sum b2 for all subsamples:
  sumb2 <- sumb2 + coef(mod2)[[2]]
}
betahat<-sumb2/C

#MSE
mean((betahat-beta)^2)

#Bias
sum(betahat-beta)

#Variance
var(betahat)
```

b.
In this question, I have to do the randomized Hadamard transform on the full data. Due to the capacity of the computer, I use the algorithm in question c.
Thus, I do the transformation in matlab.

Caculate $MSE_{\beta}$,$Bias_{\beta}$,$Variance_{\beta}$ for case1:
```{r}
#(b)
#Case 1.
hx1<-read.table("/Users/shaoyuming/x1.dat", header=F, sep = ",")
hy1<-read.table("/Users/shaoyuming/y1.dat", header=F, sep = ",") 

hzz <- cbind(hx1, hy1)
hzz<-data.frame(hzz)

N <- nrow(hx1)
C <- 1000 
S <- 1000
sumb2 <- rep(0,50)
for (i in 1:C){   # a loop over the number of subsamples
  set.seed(3*i)   # a different seed for each subsample  
  subsample <- hzz[sample(1:N, size=S, replace=TRUE), ]
  mod2 <- lm(V1.1~V1:V50,data=subsample)
  #sum b2 for all subsamples:
  sumb2 <- sumb2 + coef(mod2)[[2]]
}
hbetahat<-sumb2/C

#MSE
mean((hbetahat-beta)^2)

#Bias
sum(hbetahat-beta)

#Variance
var(hbetahat)
```

Caculate $MSE_{\beta}$,$Bias_{\beta}$,$Variance_{\beta}$ for case2:
```{r}
hx2<-read.table("/Users/shaoyuming/x2.dat", header=F, sep = ",")
hy2<-read.table("/Users/shaoyuming/y2.dat", header=F, sep = ",") 

hzz2 <- cbind(hx2, hy2)
hzz2<-data.frame(hzz2)


N <- nrow(x2)
C <- 1000 
S <- 1000
sumb2 <- rep(0,50)
for (i in 1:C){   # a loop over the number of subsamples
  set.seed(3*i)   # a different seed for each subsample  
  subsample <- hzz2[sample(1:N, size=S, replace=TRUE), ]
  mod2 <- lm(V1.1~V1:V50,data=subsample)
  #sum b2 for all subsamples:
  sumb2 <- sumb2 + coef(mod2)[[2]]
}
betahat<-sumb2/C

#MSE
mean((betahat-beta)^2)

#Bias
sum(betahat-beta)

#Variance
var(betahat)
````

